---
title: "Task 4 - Naples Reef purple urchins (McManus)"
author: "Nick McManus"
date: "10/23/2021"
output: html_document
---

# Overview

The "control" treatment did not remove any giant kelp (*Macrocystis pyriferais*) within the 2000 m$^2$ plot. The "Annual" treatment involved yearly removal of giant kelp at an adjacent 2000 m$^2$ plot and allowed to recolonize throughout the year. 



```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
# attach packages
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)
```

```{r}
# read in data
urchins <- read_csv(here("data", "urchins.csv"))

# clean data and make columns easier to work with in next step
urchins_clean <- clean_names(urchins) %>% 
  rename(size_cm = size) %>% 
  mutate(date = mdy(date),
         year = year(date))
```


Filter the cleaned data to keep only observations for purple sea urchins at NAPL in 2010.
```{r}
purple_urchins_napl_2010 <- urchins_clean %>% 
  # filter for only only Naples Reef, purple urchins, and the year 2010
  filter(site == "NAPL",
         year == "2010",
         common_name == "Purple Urchin")
```

Create a subset containing helpful statics, then display in a summary table. 
```{r}
 # compare the mean size between different treatment types
  purple_urchins_stats <- purple_urchins_napl_2010 %>% 
    group_by(treatment) %>% 
    summarize(
          mean_size_cm = round(mean(size_cm), digits = 1),
          median_size_cm = round(median(size_cm), digits = 1),
          sd_size_cm = round(sd(size_cm), digits = 1),
          sample_size = n())


# return sample statistic values in table format
purple_urchins_stats %>% 
  kable(col.names = c("Treatment type",
                      "Mean (cm)",
                      "Median (cm)",
                      "Standard deviation (cm)",
                      "Sample size (n)"),
        caption = "**Table 1.** Comparison of purple urchin sizes between treatment types at Naples Reef in 2010")  %>% 
  
  kable_styling(full_width = FALSE,
                bootstrap_options = "hover",
                position = "left")
  
```


```{r}
ggplot(data = purple_urchins_napl_2010, aes(x = size_cm)) +
  geom_histogram() +
  facet_wrap(~ treatment)
```


```{r}
ggplot(data = purple_urchins_napl_2010, aes(sample = size_cm)) +
  geom_qq() +
  facet_wrap(~ treatment)
```

