---
title: "Task 3 - Santa Barbara urchins (McManus)"
author: "Nick McManus"
date: "10/22/2021"
output: 
  html_document: 
    toc: yes
    theme: spacelab
---

# [Overview]{.ul}

In this task, I work with a dataset for purple (*Strongylocentrotus purpuratus*) and red (*Mesocentrotus franciscanus*) sea urchins observed at five locations within the Santa Barbara Channel Long Term Ecological Research (SBC LTER) site. Specifically, this study focused on the effect of kelp removal on size frequency distribution of urchins. The diameter of urchins was measured to the nearest 0.5cm within 40m x 2m sampling plots. For this task, I first explore the difference in size distribution between urchin species, then examine how purple urchin counts between sites have changed over time. 

**Data citation:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. <https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b>.


# [Setup]{.ul}

### Attach Packages

First I need to attach all packages used for this analysis.

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# attach packages
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)
```

### Data Prep

Here I read in the raw data from `urchins.csv` and return the first 6 lines for reference.

```{r}
#read in data
urchins <- read_csv(here("data","urchins.csv"))

#return first 6 lines
head(urchins)
```

Next, I will clean the data to make the data easier to work with.

```{r}
#convert column names to lowersnake case
urchins_clean <- clean_names(urchins) %>% 
  
  #add units to size column for clarity
  rename(size_cm = size) %>% 
  
  #change date to standardized format
  mutate(date = mdy(date))
  
```


# [Urchin Size Distribution by Species]{.ul}

In this section I compare the distribution of sizes (in cm) between purple and red urchins. Using the cleaned data, I visually present the differences in distribution with a histogram (Fig. 1) and summary statistics table (Table 1).

```{r, fig.cap = "**Fig. 1: Distribution of purple and red sea urchin sizes found at SBC LTER research sites.** Purple urchins (*S. purpuratus*, left) and red urchins (*M. franciscanus*, right) were measured to the 0.5cm at five sites within the SBC LTER. The purple urchins sample has a mean of 4.5cm and std dev of 1.2cm, while the red urchin sample has a mean of 6.8cm and std dev of 1.9cm."}

ggplot(data = urchins_clean) +
  # create histogram and set aesthetics
  geom_histogram(
    aes(x = size_cm, fill = common_name),
    color = "black",
    bins = 12,
    show.legend = FALSE) +  #remove redundant legend
  
  #change axis labels
  labs(x = "Urchin diameter (cm)",
       y = "Frequency") +
  
  #set fill colors
  scale_fill_manual(values = c("darkorchid", "firebrick2")) +
  
  #make histogram faceted to urchin type 
  facet_wrap(~ common_name) +
  
  #change text elements and set theme
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)) +
  
  theme_bw()

```


```{r}
# create data subset containing urchin stats of interest
urchins_clean_stats <- urchins_clean %>% 
  group_by(common_name) %>% 
  summarize(mean_size_cm = round(mean(size_cm), digits = 1),
            median_size_cm = round(median(size_cm), digits = 1),
            sd_size_cm = round(sd(size_cm), digits = 1),
            sample_size = n())


# create table to display stats
urchins_clean_stats %>% 
  kable(col.names = c("Species",
                      "Mean (cm)",
                      "Median (cm)",
                      "Standard deviation (cm)",
                      "Sample size (n)"),
        caption = "**Table 1.** Descriptive statistics of purple and red urchin sizes.") %>% 
  
  #customize table styling
  kable_classic(full_width = FALSE,
                html_font = "Cambria",
                position = "left")
```

The sample of purple urchins have a mean and median of 4.5cm, while red urchins have a mean of 6.8cm and median of 7.0cm (Table 1). Both the mean and median appear to accurately represent the central tendency of urchin sizes. The distribution for purple urchin sizes is fairly normal, with a slight positive skew (Figure 1), and a skews positively and has a smaller spread (std dev). Red urchin sizes are spread more and normally distributed.



# [Purple urchin counts by site and year.]{.ul}

In this section I focus on purple urchins and explore the changes in count by research site over time. To do this, I first create a dataframe containing only observations for purple urchins. A column with full research site names is added for clarification.

```{r}
#create subset of data for purple urchins
purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = ymd(date),
         year = year(date),        #new column with year
         site_full = case_when(    #change site abbreviations to full names
           site == "NAPL" ~ "Naples Reef",
           site == "AQUE" ~ "Arroyo Quemado Reef",
           site == "CARP" ~ "Carpinteria Reef",
           site == "MOHK" ~ "Mohawk Reef",
           site == "IVEE" ~ "Isla Vista Reef"))
```

Next, I create a subset grouping site and year and returning the sample size.

```{r}
# find total number of purple urchins observed by site and year
purple_urchins_site_year <- purple_urchins %>% 
  group_by(site_full, year) %>% 
  summarize(sample_size = n())

head(purple_urchins_site_year)
```


Finally, I visually present the changes of purple urchin count over time utilizing a line graph.

```{r, fig.cap = "**Figure 2. Change in Purple Urchin counts between research locations from 2008 to 2020.** Different research sites are represented by different colored lines in the graph. From to 2008 and 2011, Purple urchins were counted at four different research sites. Isla Vista Reef began in 2012."}
ggplot(data = purple_urchins_site_year, aes(x = year, y = sample_size)) +
  
  geom_line(aes(color = site_full),
            size = 1.3,      #change line width
            alpha = 0.9) +   #change line transparency
  
  #change line colors
  scale_color_manual(values = c("magenta", "darkblue", "aquamarine4", "firebrick3", "goldenrod")) +
  
  #customize axis tick marks
  scale_x_continuous(breaks = 2008:2020) +
  scale_y_continuous(breaks = seq(min(0), max(1500), by = 250)) +
  
  #change labels
  labs(y = "Urchin count",
       x = "Year",
       color = "Research Site") +
  
  #set theme
  theme_classic() + 
  
  #customize legend position
  theme(legend.position = c(0.85, 0.8),
        legend.title = element_text(face = "bold"),
        axis.title.x = element_text(face = "bold"), 
        axis.title.y = element_text(face = "bold")) +
  
  #customize legend title position. Unnecessary here, but included for future reference!
  guides(color = guide_legend(title.position = "top"))
```
Assume that the methodology (surveyed area size) remains constant every where. 
