---
title: "Task 4 - Naples Reef purple urchins (McManus)"
author: "Nick McManus"
date: "10/23/2021"
output: 
  html_document: 
    toc: yes
    theme: spacelab
---

# Overview

This task utilizes the same urchin dataset as Task 3. In this task, I explore the difference between two experimental treatment types. The "control" treatment did not remove any giant kelp (*Macrocystis pyriferais*) within the 2000 m$^2$ plot. The "Annual" treatment involved yearly removal of giant kelp at an adjacent 2000 m$^2$ plot and allowed to recolonize throughout the year. Utilizing the `urchins.csv` data, I see how these two treatments affected size distribution of purple urchins at Naples Reef in 2010.

**Data citation:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. <https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b>.




# Setup

### Attach packages

First I need to attach all packages used for this analysis.

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# attach packages
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)
```

### Data prep

Here I read in the raw data from `urchins.csv`, then clean the data to make the data easier to work with. Afterward I return the first 6 lines for reference.

```{r}
# read in data
urchins <- read_csv(here("data", "urchins.csv"))

# clean data and make columns easier to work with in next step
urchins_clean <- clean_names(urchins) %>% 
  rename(size_cm = size) %>% 
  mutate(date = mdy(date),
         year = year(date))

head(urchins_clean)
```



# Data wrangling

Filter the cleaned data to keep only observations for purple sea urchins at NAPL in 2010.

```{r}
purple_urchins_napl_2010 <- urchins_clean %>% 
  # filter for only only Naples Reef, purple urchins, and the year 2010
  filter(site == "NAPL",
         year == "2010",
         common_name == "Purple Urchin")
```

Create a subset containing the descriptive statics.

```{r}
 # compare the mean size between different treatment types
  purple_urchins_stats <- purple_urchins_napl_2010 %>% 
    group_by(treatment) %>% 
    summarize(
          mean_size_cm = round(mean(size_cm), digits = 1),
          median_size_cm = round(median(size_cm), digits = 1),
          sd_size_cm = round(sd(size_cm), digits = 1),
          sample_size = n())
```



# Data visualization

Here I use `purple_urchins_napl_2010` to create exploratory faceted histogram and QQ-plot to observe the size distribution of purple urchins.

```{r}
# create histogram
ggplot(data = purple_urchins_napl_2010, aes(x = size_cm)) +
  geom_histogram() +
  facet_wrap(~ treatment)
```

```{r}
#create qqplot
ggplot(data = purple_urchins_napl_2010, aes(sample = size_cm)) +
  geom_qq() +
  facet_wrap(~ treatment)
```

Based off the QQ-plot alone, both treatments appear to be equally distributed. Looking at the histograms, it appears that the Annual treatment has a slight positive skew. Taking both figures into consideration, distribution and spread of data between treatments seems fairly equal with a slightly different central tendency. 


Next I use the subset `purple_urchins_stats` to create a summary table.
```{r}
# return descriptive statistic values in table format
purple_urchins_stats %>% 
  kable(col.names = c("Treatment type",
                      "Mean (cm)",
                      "Median (cm)",
                      "Standard deviation (cm)",
                      "Sample size (n)"),
        caption = "**Table 1.** Descriptive statistics of purple urchin sizes between treatment types at Naples Reef in 2010")  %>% 
  
  kable_styling(full_width = FALSE,
                bootstrap_options = "hover",
                html_font = "Cambria",
                position = "left")
```

Table 1 confirms that the spread of data between treatment types is equal and there is a 5% difference (0.2cm) in means values. 
